\Dune, the Distributed and Unified Numerics Environment is a modular toolbox for solving partial differential equations with grid-based methods. It supports easy discretization using methods like Finite Elements, Finite Volumes, and also Finite Differences.

\Dune\ is free software licensed under the GPL with a so called "runtime exception" (\url{http://www.dune-project.org/license.html}).

The main intention is to create slim interfaces allowing an efficient use of legacy and/or new libraries. Using C++ techniques \Dune\ allows to use very different implementations of the same concept (i.e. grids, solvers, ...) using a common interface with a very low overhead. More: \url{http://www.dune-project.org/dune.html} \cite{duneweb}.



\section{Core Modules}
The framework consists of a number of modules which are downloadable as separate packages. The current core modules are:

\begin{itemize}
 \item \texttt{dune-common}\\
    contains the basic classes used by all \Dune-modules. It provides some infrastructural classes for debugging and exception
    handling as well as a library to handle dense matrices and vectors.
 \item \Grid\\
    is the most mature module. It defines nonconforming, hierarchically nested, multi-element-type, parallel grids in
    arbitrary space dimensions. Graphical output with several packages is available, e.g. file output to IBM data explorer
    and VTK (parallel XML format for unstructured grids). The graphics package Grape has been integrated in interactive mode.
 \item \texttt{dune-istl} (Iterative Solver Template Library)\\
    provides generic sparse matrix/vector classes and a variety of solvers based on these classes. A special feature is the use
    of templates to exploit the recursive block structure of finite element matrices at compile time. Available solvers include
    Krylov methods, (block-) incomplete decompositions and aggregation-based algebraic multigrid.
\end{itemize}




\section{Available Grid Implementations}\label{availablegrids}
 So far six grid implementations are available through the \Grid interface. Each is geared towards a different purpose.
\begin{itemize}
    \item \texttt{ONEDGRID}: A sequential locally adaptive grid in one space dimension
    \item \texttt{SGRID}: A structured grid in n space dimensions
    \item \texttt{YASPGRID}: A structured parallel grid in n space dimensions (used as default grid)
    \item \texttt{ALUGRID\_CUBE, ALUGRID\_SIMPLEX}: A parallel locally adaptive grid in 2 and 3 space dimensions
    \item \texttt{ALBERTAGRID}: The grid manager of the Alberta toolbox
    \item \texttt{UGGRID}: The grid manager of the UG toolbox (UG is not freely available)
\end{itemize}
More information about these grids: \url{http://www.dune-project.org/doc/devel/features.html}.



\input{installation_body.tex}
